---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# osmar2

osmar2 is a version of the [osmar package](http://cran.r-project.org/package=osmar) 
("OpenStreetMap and R") rewritten to use xml2 package instead of XML, 
and some other adaptations to allow for much quicker reading of .osm files. Currently, 
only the functions for data import (`get_osm()` for `osmsource_file()` 
and the functions it calls have been changed), and I believe I have done so in a
way that is backwards compatible, (or at least, it has been in the ways I have used 
osmar objects so far).

## Speed Comparison
```{r}
# install.packages("osmar")
# install.packages("devtools")
# devtools::install_github("gergness/osmar2")
# install.packages("microbecnhmark")
# install.packages("ggplot2")

suppressMessages({
  library(osmar2)
  library(microbenchmark)
  library(ggplot2)
})

tiny_file <- tempfile()
download.file("https://dl.dropboxusercontent.com/u/2019891/osmar2/TC_tiny_osm_file.osm", tiny_file)

mb_tiny <- microbenchmark(
  osmar = osmar::get_osm(complete_file(), osmsource_file(tiny_file)),
  osmar2 = osmar2::get_osm(complete_file(), osmsource_file(tiny_file)),
  times = 5
)

mb_tiny
```


Even more benefits as files get bigger
```{r} 
small_file <- tempfile()
download.file("https://dl.dropboxusercontent.com/u/2019891/osmar2/TC_small_osm_file.osm", small_file)

mb_small <- microbenchmark(
  osmar = osmar::get_osm(osmar::complete_file(), osmar::osmsource_file(small_file)),
  osmar2 = osmar2::get_osm(osmar2::complete_file(), osmar2::osmsource_file(small_file)),
  times = 1
)

mb_small

speed_df <- data.frame(
  `file size (Mb)` = rep(file.size(c(tiny_file, small_file)), each = 2) / (1012 ^ 2),
  pkg = rep(c("osmar", "osmar2"), 2),
  `time (s)` = c(summary(mb_tiny, unit = "s")$mean, 
            summary(mb_small, unit = "s")$mean), 
  check.names = FALSE
)

ggplot(data = speed_df, aes(x = `file size (Mb)`, y = `time (s)`, group = pkg, color = pkg)) +
  geom_path() + geom_point() + 
  scale_y_log10() +
  annotation_logticks(base = 10, sides = "l")
 
```

osmar2 even has tolerable load times (~7 minutes) for a ~125Mb file. I haven't
checked how long this would take on osmar.
```{r, eval = FALSE}
# Too slow to run regularly...
big_file <- tempfile()
download.file("https://dl.dropboxusercontent.com/u/2019891/osmar2/TC_big_osm_file.osm", big_file)

mb_big <- microbenchmark::microbenchmark(
 osmar2 = osmar2::get_osm(complete_file(), osmsource_file(big_file)),
 times = 1
)

mb_big
```

## Development

I don't plan to do much further development on osmar2, but wanted to post it publicly 
in case it is useful for anyone else. I believe that it could be written to be much faster
with, because it is currently scanning through the XML document many times. 

I will try to help with issues and work with PRs, but please be patient :)